---
title: "Stein"
author: "Maximilian Stein"
date: "`r Sys.Date()`"
format:
  html: default
  pdf:
    documentclass: article
    keep-tex: true
execute:
  echo: true
  warning: false
  message: false
---

# Introduction 

This document contains solutions for the exam, making use of the `dplyr`, `tidyr` and `stringr` libraries.

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(here)
```

## Question 1

Load the data.

```{r}
grades_data <- read.csv(here("grades.csv"))
```

## Question 2

Give the number of grades in the data set directly in the text of your quarto file, in a sentence of the form “The data set contains xxx grades.” where xxx is replaced by the number of grades. This must be computed from the data set.

```{r}
cat("The data set contains", nrow(grades_data), "grades.")
```

## Question 3 

Use distinct() compute the number of students on the data set and report the value directly in the text as in the previous question.

```{r}
num_students <- grades_data %>%
  distinct(id) %>%
  nrow()

cat("There are", num_students, "students in the data set.")
```

